---
title: "requesting_data"
author: "Yu Dong"
date: "11/29/2016"
output: html_document
---

# Load dataset

Load the downloaded dataset from https://data.lacity.org/A-Well-Run-City/MyLA311-Service-Request-Data-2016/ndkd-k878

```{r load}
setwd("~/Desktop/DSO545/Final project")
# load("processed1.RData")
library(dplyr)
library(ggplot2)
library(lubridate)
library(stringr)
request_data = read.csv("raw data/MyLA311_Service_Request_Data_2016.csv")
```

# Preprocessing

Preprocess the data:  
change the format of time variables using lubridate package  

```{r preprocess}
request_data$CreatedDate = mdy_hms(request_data$CreatedDate)
request_data$UpdatedDate = mdy_hms(request_data$UpdatedDate)
# since there are lots of observations like xx/xx/xxxx 12:00:00AM which leads to unaccurate time records, remove the hour-minute_seconds part
request_data$ServiceDate = str_replace_all(request_data$ServiceDate, " [0-9]{2}:[0-9]{2}:[0-9]{2} [AP]M", "")
request_data$ServiceDate = mdy(request_data$ServiceDate)
# attention: there are many obviously unreasonable records: many records with service date of 1900/01/01, and some other service dates in 2017 or later
request_data$ClosedDate = mdy_hms(request_data$ClosedDate)
```

Calculate new variables:  
1. extract the month, weekday and hour the request created  
2. create the response period in hours as the time from created date to closed date

```{r new variables}
# extract the month, weekday and hour of CreatedDate
request_data$month_created = month(request_data$CreatedDate)
request_data$weekday_created = wday(request_data$CreatedDate, label = TRUE)
request_data$hour_created = hour(request_data$CreatedDate)

# calculate response period (ClosedDate - CreatedDate)
request_data$response_period = round((request_data$ClosedDate - request_data$CreatedDate) / (60 * 60),2)  # round to 2 digits, in hours

# special attention: the percentage of requests without a ClosedDate: 0.546
nrow(filter(request_data, is.na(ClosedDate)))/ nrow(request_data)
```

# Trial run for sampled data

Sampling 1000 observations from the whole dataset

```{r sampling}
set.seed(1)
sampled_data = sample_n(request_data, 1000)
```


